### 2014-12-01 README.Rmd
update structure

---
title: "decompr article"
author: "Bastiaan Quast and Victor Kümmritz"
date: "Wednesday, November 26, 2014"
output: md_document
bibliography: bibliography.bib
---

```{r load-package, echo=FALSE}
library(decompr)
```

Abstract
===============
The R package **decompr** implements algorithms for the analysis of Global Value Chains.
Two methods for gross export flow decomposition using Inter-Country Input-Output tables are provided.
The first method is the **Wang-Wei-Zhu** [-@wang2014quantifying] algorithm which splits bilateral gross exports into 16 value added components.
These components can broadly be divided into domestic and foreign value added.
The second method concerns a **source decomposition** algorithm.
This derives the value added origin of a country-industry’s exports by source country and source industry,
by applying the basic Leontief insight to gross trade data.
This article summarises the methodology of the algorithms and describes the format of the input and output data.

Introduction
================
Global Value Chains (GVCs) have become a central factor in trade and development policy.
Policy makers from different countries and institutions have placed them at the center of their agenda
and continuously emphasize their growing importance for both international trade and economic development.
However, analysing this phenomenon empirically requires complex matrix manipulations,
since the relevant data is only available in the form of gross flows.
This package enables researchers with little background in matrix algebra and linear programming
to easily derive standard GVC indicators for statistical analysis.

Data
=================
A sample data set of Input-Output tables is included.
In order to save space and speed up computations, the included data set uses regional aggregates.

Load the included data set of WIOD regional Input-Output tables.
```{r load-data}
data(wiod)
```

A key step in loading the data is making sure that the dimensions of the data inputs are correct,
in the sample data we have **G = 7** (generally countries, here regions) and **N = 43** (number of industries).
```{r data-dimensions}
dim(intermediate_demand) # (2 + GN + totals) x (2 + GN)
dim(final_demand)        # (2 + GN + totals) x (2 + G*5)
```

Note thata **G** and **N** are automatically inferred from the dimensions of the data.

When importing data, be sure to specify **stringsAsFactors = FALSE** and to **header=FALSE**,
as the Input-Output tables use two header lines (country and industry),

Methods
=================
 * Small derivation
 * R demonstration using WIOD data

```{r load_tables}
elements <- load_tables( x = intermediate_demand,
                         y = final_demand         )
```

Wang Wei Zhu decomposition
------------------------------
We can now decompose the elements using the Wang-Wei-Zhu (or source) decomposition.
```{r wwz}
wwz <- wwz( elements )
```

In addition, a wrapper function called **decomp** is provided which performs both steps at once.
Though it is recomended that the seperate functions be used for large data sets.

```{r wwz-wrapper}
wwz2 <- decomp( x = intermediate_demand,
                y = final_demand,
                method = "wwz"           )
```

Note that **wwz** is the default method.

The output data is as follows

```{r wwz-eval}
str(wwz)
```

Source decomposition
---------------------------------------
 * Small derivation
 * R demonstration using WIOD data

We can now decompose the elements using the Source decomposition.
```{r source}
source <- kung_fu( elements )
```

In addition, a wrapper function called **decomp** is provided which performs both steps at once.
Though it is recomended that the atomic functions be used for large data sets.

```{r source-wrapper}
source2 <- decomp( x = intermediate_demand,
                y = final_demand,
                method = "source"           )
```

The output data is as follows

```{r source-eval}
str(source)
```

Discussion
------------------------

Acknowledgements
--------------------
FNS
Fei Wang

References
----------------------